Feeding Dynamic Behavior into Optimization: Simulation-based Multi-objective Overtime Planning

### 2014-09-03 data/result/Article/Latex/Article.tex
.

% Table generated by Excel2LaTeX from sheet 'tabela 3'
\begin{table}[htbp]
  \scriptsize
	\centering
  \caption{Effect-size for comparisons between NSGAII and each OMS in terms of each quality indicator.}
    \begin{tabular}{p{0.6cm} p{0.6cm} p{0.65cm} p{0.65cm}p{0.65cm}p{0.65cm}p{0.65cm}p{0.65cm}}
    \toprule
    \textbf{QI} & \textbf{OMS} & \textbf{ACAD} & \textbf{WMET} & \textbf{WAMS} & \textbf{PSOA} & \textbf{OMET} & \textbf{PARM} \\
    \midrule
    \multirow{3}[1]{*}{I$_{C}$} 
		      & MAR   & 94\%  & 89\%  & 96\%  & 85\%  & 80\%  & 44\% \\
          & SH    & 84\%  & 81\%  & 88\%  & 85\%  & 60\%  &  7\% \\
          & CPM   & 94\%  & 89\%  & 96\%  & 85\%  & 80\%  & 44\% \\\hline
    \multirow{3}[2]{*}{I$_{HV}$} 
		      & MAR   & 100\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
          & SH    & 100\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
          & CPM   &   n/a & 100\% & 100\% &   n/a & 100\% &   n/a \\\hline
    \multirow{3}[2]{*}{I$_{GD}$} 
		      & MAR   & 100\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
          & SH    & 100\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
          & CPM   &   n/a & 100\% & 100\% &   n/a & 100\% &   n/a \\
    \bottomrule
    \end{tabular}%
\end{table}%

@@

% Table generated by Excel2LaTeX from sheet 'tabela 4'
\begin{table}[htbp]
	\scriptsize
  \centering
  \caption{Quality indicators for each instance under our approach (NSGAII) and a model without error dynamics (NSGA$_{NE}$). Pareto fronts for NSGA$_{NE}$ cover a larger part of the solution space, leading to smaller $I_{GD}$. Solutions produced under NSGAII tend to concentrate in certain parts of the solution space}
    \begin{tabular}{p{0.8cm}p{1.0cm}ccc}
    \toprule
    \textbf{Instance} & \textbf{Config.} & \textbf{I$_{C}$} & \textbf{I$_{HV}$} & \textbf{I$_{GD}$} \\
    \midrule
    \multirow{2}[2]{*}{ACAD} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.530 $\pm$ 0.006} &         0.004 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.004  &         0.333 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.00}  \\\hline
    \multirow{2}[2]{*}{WMET} 
				  & NSGAII      &         0.020 $\pm$ 0.012  & \textbf{0.544 $\pm$ 0.004} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.003  &         0.334 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{WAMS} 
				  & NSGAII      &         0.020 $\pm$ 0.010  & \textbf{0.498 $\pm$ 0.003} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ & \textbf{0.020 $\pm$ 0.003} &         0.337 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{PSOA} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.471 $\pm$ 0.006} &         0.004 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.002  &         0.338 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{OMET} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.523 $\pm$ 0.004} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.003  &         0.337 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{PARM} 
				  & NSGAII      &         0.020 $\pm$ 0.018  & \textbf{0.462 $\pm$ 0.007} &         0.006 $\pm$ 0.001  \\
          & NSGA$_{NE}$ & \textbf{0.020 $\pm$ 0.003} &         0.338 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\
    \bottomrule
    \end{tabular}%
\end{table}%

@@

{\^A12} effect-sizes for each quality indicator are presented in Table V. It can be observed that NSGAII consistently presents higher I$_{HV}$ (good) and higher I$_{GD}$ (bad) compared to NSGA$_{NE}$. On regard of I$_{C}$, the algorithms are competitive, with small effect-sizes favoring NSGA$_{NE}$. These results are explained by the larger number of solutions produced by NSGA$_{NE}$: its Pareto fronts have on average 1.8 times the number of solutions of fronts calculated for NSGAII, reaching up to 3.9 times this number for the largest instance (PARM). The larger number of solutions allows NSGA$_{NE}$ to cover parts of the reference front uncovered by NSGAII, lowering its I$_{GD}$ and increasing its I$_{C}$. NSGAII contributes with more points than NSGA$_{NE}$ on parts of the reference front that it covers, but these improvements are not sufficient to compensate for the distance between points in the extremes of its fronts and the parts of the reference front defined only by NSGA$_{NE}$.

@@

Besides quality indicators, it is interesting to consider the differences observed in objective values produced under the formulations with and without overtime error dynamics. Table VI presents the differences in average makespan and cost observed in solutions involving the highest overtime allocation for each instance (about 3 hours of daily overtime over the project life-cycle). Data shows that NSGA$_{NE}$ tends to underestimate project cost and makespan when a large number of overtime hours are imposed to the development team. Such difference, observed in 5 out of 6 instances, is due to longer testing activities required to identify and correct the exceeding errors introduced in the software by tired developers.

>>>>

% Table generated by Excel2LaTeX from sheet 'tabela 3'
\begin{table}[htbp]
  \scriptsize
	\centering
  \caption{Effect-size for comparisons between NSGAII and each OMS in terms of each quality indicator.}
    \begin{tabular}{llcccccc}
    \toprule
    \textbf{QI} & \textbf{OMS} & \textbf{ACAD} & \textbf{WMET} & \textbf{WAMS} & \textbf{PSOA} & \textbf{OMET} & \textbf{PARM} \\
    \midrule
    \multirow{3}[1]{*}{I$_{C}$} 
		      & MAR   & 94\%  & 89\%  & 96\%  & 85\%  & 80\%  & 44\% \\
          & SH    & 84\%  & 81\%  & 88\%  & 85\%  & 60\%  &  7\% \\
          & CPM   & 94\%  & 89\%  & 96\%  & 85\%  & 80\%  & 44\% \\\hline
    \multirow{3}[2]{*}{I$_{HV}$} 
		      & MAR   & 100\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
          & SH    & 100\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
          & CPM   &   n/a & 100\% & 100\% &   n/a & 100\% &   n/a \\\hline
    \multirow{3}[2]{*}{I$_{GD}$} 
		      & MAR   & 100\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
          & SH    & 100\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
          & CPM   &   n/a & 100\% & 100\% &   n/a & 100\% &   n/a \\
    \bottomrule
    \end{tabular}%
\end{table}%

@@

% Table generated by Excel2LaTeX from sheet 'tabela 4'
\begin{table}[htbp]
	\scriptsize
  \centering
  \caption{Quality indicators for each instance under our approach (NSGAII) and a model without error dynamics (NSGA$_{NE}$). Pareto fronts for NSGA$_{NE}$ cover a larger part of the solution space, leading to smaller $I_{GD}$. Solutions produced under NSGAII tend to concentrate in certain parts of the solution space}
    \begin{tabular}{llccc}
    \toprule
    \textbf{Instance} & \textbf{Config.} & \textbf{I$_{C}$} & \textbf{I$_{HV}$} & \textbf{I$_{GD}$} \\
    \midrule
    \multirow{2}[2]{*}{ACAD} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.530 $\pm$ 0.006} &         0.004 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.004  &         0.333 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.00}  \\\hline
    \multirow{2}[2]{*}{WMET} 
				  & NSGAII      &         0.020 $\pm$ 0.012  & \textbf{0.544 $\pm$ 0.004} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.003  &         0.334 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{WAMS} 
				  & NSGAII      &         0.020 $\pm$ 0.010  & \textbf{0.498 $\pm$ 0.003} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ & \textbf{0.020 $\pm$ 0.003} &         0.337 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{PSOA} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.471 $\pm$ 0.006} &         0.004 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.002  &         0.338 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{OMET} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.523 $\pm$ 0.004} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.003  &         0.337 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{PARM} 
				  & NSGAII      &         0.020 $\pm$ 0.018  & \textbf{0.462 $\pm$ 0.007} &         0.006 $\pm$ 0.001  \\
          & NSGA$_{NE}$ & \textbf{0.020 $\pm$ 0.003} &         0.338 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\
    \bottomrule
    \end{tabular}%
\end{table}%

@@

{\^A12} effect-sizes for each quality indicator are presented in Table V. It can be observed that NSGAII consistently presents higher I$_{HV}$ (good) and higher I$_{GD}$ (bad) compared to NSGA$_{NE}$. On regard of I$_{C}$, the algorithms are competitive, with small effect-sizes favoring NSGA$_{NE}$. These results are explained by the larger number of solutions produced by NSGA$_{NE}$: its Pareto fronts have on average 1.4 times the number of solutions of fronts calculated for NSGAII, reaching up to 2.7 times this number for the largest instance (PARM). The larger number of solutions allows NSGA$_{NE}$ to cover parts of the reference front uncovered by NSGAII, lowering its I$_{GD}$ and increasing its I$_{C}$. NSGAII contributes with more points than NSGA$_{NE}$ on parts of the reference front that it covers, but these improvements are not sufficient to compensate for the distance between points in the extremes of its fronts and the parts of the reference front defined only by NSGA$_{NE}$.

@@

Besides quality indicators, it is interesting to consider the differences observed in objective values produced under the formulations with and without overtime error dynamics. Table VI presents the differences in average makespan and cost observed in solutions involving the highest overtime allocation for each instance (about 3 hours of daily overtime over the project life-cycle). Data shows that NSGA$_{NE}$ underestimates project cost and makespan when a large number of overtime hours are imposed to the development team. Such difference, observed for all instances, is due to longer testing activities required to identify and correct the exceeding errors introduced in the software by tired developers.
